---
title: "Thoughts on America's Economy"
author: "Victor Bowker, Harvard College"
description: "Gov 50 Final Project, Fall Semester 2023"
output:
  distill::distill_article:
    self_contained: false
---

---
output:
  pdf_document: default
  html_document: default
---

```{r}
##beginning, loading data and assigning names
library(tidyverse)
library(haven)
altdat <- read_sav("W52_Jul19/ATP W52.sav")
dat <- read_sav("W110_Jun22/ATP W110.sav")
```

Introduction
---

 How do Americans feel about the economy? Does their party, gender, or age impact how they feel? What about when the President they support is in office or isn't? I hypothesize that in general, when a President representing Party A is in the White House, voters in Party A will feel positively about the economy. Similarly, when the President represents Party A, members of Party B will feel quite negatively. I hypothesize this because the economic positions of political parties vary greatly, meaning that the President will often make substantial changes to align the economy with policies similar to their party. In our hyper-polarized political climate, Americans cling to one side, generally only supporting one party. This leads to further division, and economic changes each time a new President enters the Oval Office. 

  My data is compiled from surveys of Americans in 2019 and 2022, representing President Trump's and President Biden's first terms, respectively. Similar questions were asked in each survey, regarding things such as abortion, presidential approval, climate change, and the economy. This dataset also includes important information to identify the demographics of the participants, including age, gender, location in the nation, and political identity. 

  My explanatory variable is 'Party', with complementing and adjacent variables 'Age' and 'Gender'. With this variable, I will work to find my outcome variable, the feelings towards the economy, represented with `ECON1_W110` and `ECON1_W52`, representing current economic feelings under President Biden and President Trump respectively. I am also utilizing `ECON1B_W110` and `ECON1B_W52` to represent feelings toward the future of the American economy, under President Biden and President Trump respectively. 
  
  It is important to note that at the time of this project, neither President Joe Biden nor President Donald Trump had served a second term in office. The data will refer to information resulting from their first term, as that is shared common ground. 

Data
---

  For this project, I utilized freely accessible data from the Pew Research Center American Trends Panel, which is composed of responses from over 10,000 American adults. The Research Center maintains ties with these 10,000, working to compile results from the same group Pew strives to find a diverse group of respondents, weighing to align with national demographics. The specific data sets, ATP W52, from July 2019, and ATP W110, from June 2022, represent benchmarks in each President's first term of office. Pew's ATP W52 contains values from 4,175 unique respondents, while ATP W110 contains data from 6,174 entries. The makeup of each political party and gender is demonstrated below. 
  
```{r}
##mutating Biden data to simplify label names
var <- 
dat |>
mutate(
    Party = if_else(PARTY_W110 == 1, "Republican", "Democrat"), 
    Approval = if_else(POL1JB_W110 == 1, "Approve", "Disapprove"),
    Gender = if_else(F_GENDER == 1, "Man", "Woman"),
    Age = if_else(F_AGECAT == 1, "18-29",
    if_else(F_AGECAT == 2, "30-49",
    if_else(F_AGECAT == 3, "50-64",
    if_else(F_AGECAT == 4, "65+", "N/A")))))
```
```{r}
##mutating Trump data to simplify label names
var2 <- 
altdat |>
mutate(
    Party = if_else(F_PARTY_FINAL == 1, "Republican", "Democrat"), 
    Gender = if_else(F_SEX == 1, "Man", "Woman"),
    Approval = if_else(POL1DT_W52 == 1, "Approve", "Disapprove"),
    Age = if_else(F_AGECAT == 1, "18-29",
    if_else(F_AGECAT == 2, "30-49",
    if_else(F_AGECAT == 3, "50-64",
    if_else(F_AGECAT == 4, "65+", "N/A")))))
```
```{r}
##demonstrating count of respondents for each dataset
modelcount <-
  altdat |>
  nrow()

modelcount2 <-
  dat |>
  nrow()
```
```{r}
##demonstrating political makeup of respondents for each dataset
political <-
  var |>
  group_by(Party) |>
  summarize(Count = n())

political2 <-
  var2 |>
  group_by(Party) |>
  summarize(Count = n())
```
```{r}
##demonstrating gender of respondents for each dataset
gendermak <-
  var |>
  group_by(Gender) |>
  summarize(Count = n())

gendermak2 <-
  var2 |>
  group_by(Gender) |>
  summarize(Count = n())

##printing count, party, and gender of respondents in data sets

##trump
knitr::kable(modelcount2, col.names = c("Count"),
             digits = 2,
             caption = "Number of Respondents in ATP W52")
knitr::kable(political2, col.names = c("Party", "Count"),
             caption = "Political Makeup of W52")
knitr::kable(gendermak2, col.names = c("Gender", "Count"),
             caption = "Gender Makeup of W52")
##biden
knitr::kable(modelcount, col.names = c("Count"),
             digits = 2,
             caption = "Number of Respondents in ATP W110")
knitr::kable(political, col.names = c("Party", "Count"),
             caption = "Political Makeup of W110")
knitr::kable(gendermak, col.names = c("Gender", "Count"),
             caption = "Gender Makeup of W110")

```

  
  
  The American Trends Panel asks generic questions to compile demographic information, as well as specific information on current topics. I identified the two data sets ATP W52 and ATP W110 because they both ask specifically about the economy. The first question is "Thinking about the nation’s economy...How would you rate economic conditions in this country today?" with possible responses of "Excellent", "Good", "Only Fair", or "Poor". This very question was posed in both data sets. Additionally, each set asked "A year from now, do you expect that economic conditions in the country as a whole will be..." with options "Better", "Worse" and "About the same as now"
  


Setting the Stage ~ Biden Approval Rating
---

From Pew Research Center data collected in 2022, President Joe Biden's approval ratings are reported below. You will find in the primary table that, serving as little surprise to many, nearly 100% of poll respondents who identify as Republicans disapproved of the President. In less resounding results, Democrats were nearly split 50/50, with a slight edge in the disapproval column for men, and a slight approval win for women. 

This data is important to include in the report as it alludes to the general satisfaction with the President at the time of the polling. In the case where all polled Americans disapproved of Biden's work, it would be possible that the data on economic stability would be less a sign of the economy and more a sign of general disliking for President Biden. This data does demonstrate that Republicans were against Biden, and the Democrats were split. 

```{r echo = FALSE}
##mutating to simplify label names
var <- 
  dat |>
  mutate(
  Party = if_else(PARTY_W110 == 1, "Republican", "Democrat"), 
  Approval = if_else(POL1JB_W110 == 1, "Approve", "Disapprove"),
  Gender = if_else(F_GENDER == 1, "Man", "Woman"),
  Age = if_else(F_AGECAT == 1, "18-29",
  if_else(F_AGECAT == 2, "30-49",
  if_else(F_AGECAT == 3, "50-64",
  if_else(F_AGECAT == 4, "65+", "N/A")))))
```
```{r}
##assigning vars and demonstrating Biden approval 
bidenvar <-
  var |>
  group_by(Party, Gender) |>
  count(Approval) |>
  mutate(pct = n / sum(n) *100 ) |>
  select(Party, Gender, Approval, pct)

##graphing Biden approval 
bidenvar |>
  ggplot(mapping = aes(x = Approval, y = pct, fill = Party)) +
  geom_col(position = "dodge") +
  facet_grid(rows = vars(Gender)) +
  scale_fill_manual(values = c(Democrat = "steelblue", Republican = "indianred")) +
  labs(title = "Approval Ratings of President Joe Biden",
  x = "Thoughts on the President",
  y = "Percentage",
  subtitle = "POL1JB_W110, ATP W110, Pew Research Center")

knitr::kable(bidenvar, col.names = c("Political Party", "Gender", "Approval Status", "Percent"),
             digits = 2,
             caption = "Approval Rating of President Joe Biden")
```



Perspectives on the Economy Under Biden
---

Moving into the substantive information, the graphs below demonstrate the results of economic stability questions, posed to respondents during President Biden's first term. Over 90% of Republicans, both male and female, reported that they felt the economy was currently "Only Fair" or "Poor". Democrats primarily identified the economy as currently "Only Fair" but still over 60% of respondents rated the economic climate in the bottom half of responses, based on positivity. 


```{r}
##filtering and assigning data for Biden current econ perspectives 
var_biden <-
  var |>
  filter(ECON1_W110 == 1 | ECON1_W110 == 2 | ECON1_W110 == 3 | ECON1_W110 == 4 | ECON1_W110 == 5)|>
  mutate(ECON1_W110 = if_else(ECON1_W110 == 1, "Excellent",
  if_else(ECON1_W110 == 2, "Good",
  if_else(ECON1_W110 == 3, "Only Fair",
  if_else(ECON1_W110 == 4, "Poor", "No Change"))))) |>
  group_by(Party, Gender) |>
  count(ECON1_W110) |>
  mutate(pct = n / sum(n) *100 ) |>
  select(Party, Gender, ECON1_W110, pct)

##assigning linear regression on Biden current perspectives 
model_biden <- lm(pct ~ ECON1_W110, data = var_biden)

##filtering and assigning data for Biden future econ perspectives
var_biden2 <-
  var |>
  filter(ECON1B_W110 == 1 | ECON1B_W110 == 2 | ECON1B_W110 == 3 | ECON1B_W110 == 4) |>
  mutate(ECON1B_W110 = if_else(ECON1B_W110 == 1, "Better",
  if_else(ECON1B_W110 == 2, "Worse",          
  if_else(ECON1B_W110 == 3, "About Same", "Unsure")))) |>
  group_by(Party, Gender) |>
  count(ECON1B_W110) |>
  mutate(pct = n / sum(n) *100 ) |>
  select(Party, Gender, ECON1B_W110, pct)

##assigning linear regression on Biden future perspectives
model_biden2 <- lm(pct ~ ECON1B_W110, data = var_biden2)

##plotting data of Biden current econ perspectives
biden_plot1 <- 
  var_biden |>
  ggplot(mapping = aes(x = ECON1_W110, y = pct, fill = Party)) +
  geom_col(position = "dodge") +
  facet_grid(rows = vars(Gender)) +
  scale_fill_manual(values = c(Democrat = "steelblue", Republican = "indianred")) +
  labs(title = "Thoughts on Current Economy Under Biden",
  x = "Feelings Towards Economy",
  y = "Percentage",
  subtitle = "ECON1_W110, ATP W110, Pew Research Center") 
  
##plotting data of Biden future econ perspectives
biden_plot2 <- 
  var_biden2 |>
  ggplot(mapping = aes(x = ECON1B_W110, y = pct, fill = Party)) +
  geom_col(position = "dodge") +
  facet_grid(rows = vars(Gender)) +
  scale_fill_manual(values = c(Democrat = "steelblue", Republican = "indianred")) +
  labs(title = "Thoughts on Future Economy Under Biden",
  x = "Feelings Towards Future Economy",
  y = "Percentage",
  subtitle = "ECON1B_W110, ATP W110, Pew Research Center") 

##printing plot and table for Biden current perspectives
biden_plot1
knitr::kable(var_biden, col.names = c("Party", "Gender", "Econ Feelings", "Percentage"),
             caption = "Thoughts on Current Economy Under Biden")
```

In the first linear regression model of this project, responses from `ECON1_W110`serve as the independent variable, with `pct` working as the dependent variable. First, the intercept, holding all else equal, comes at .76, demonstrating that the approval percent of President Joe Biden is quite small and stagnant. For respondents who believe the economy is good, (`ECON1_W110Good`), there is a 8.48 percentage point increase in percent approval of President Biden. For those who currently rate the economy as only fair, (`ECON1_W110OnlyFair`), there is a resounding 34.5 percent point increase in approval. Finally, the  `ECON1_W110Poor` responses demonstrate a 53.98 percent point increase. This data demonstrates that as feelings about the economy falter, approval increases. I will address the questions raised by this in the conclusion. 

```{r}
##printing table demonstrating linear regression data on Biden current perspectives
modelsummary::modelsummary(model_biden,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           caption = "Regression on Thoughts on Current Economy Under Biden")


```

When asked about speculating on the economy one year in the future, respondents fell upon party lines yet again. Republicans rallied, resulting in over 71% of Men and 69% of Women believing the economy would be worse. The Democrats represented a more split decision, with nearly 40% of both men and women speculating the economy will be worse; 30% of men and women hoping the economy will improve; and between 20% and 30% of Democrat men and women believing it will remain the same. 


```{r}
##printing plot and table for Biden future perspectives
biden_plot2
knitr::kable(var_biden2, col.names = c("Party", "Gender", "Future Econ Feelings", "Percentage"),
             caption = "Thoughts on Future Economy Under Biden")
```


In this case, `pct` serves as the dependent variable, with results from `ECON1B_W110` serving as the independent variable. First, the intercept, which equals 27.56, meaning holding all else equal, approval of President Joe Biden stands at 27.56 percent points. For those who expect the future economy to do better in a year (`ECON1B_W110Better`), there is a 9.05 percentage point decrease in approval. Respondents in `ECON1B_W110Worse` experience, on average, a 23.76 percentage point increase in approval. This data demonstrats an interesting perspective, primarily because both parties resoundingly responded that the economy will be worse in one year. `ECON1B_W110Worse` demonstrates a significant p-value with a confidence interval of .05, but `ECON1B_W110Better` does not. 

```{r}
##printing table demonstrating linear regression for Biden future perspectives
modelsummary::modelsummary(model_biden2,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           caption = "Regression on Thoughts on Future Economy Under Biden")

```

```{r echo = FALSE}
##mutating to simplify label names
var2 <- 
  altdat |>
mutate(
    Party = if_else(F_PARTY_FINAL == 1, "Republican", "Democrat"), 
    Gender = if_else(F_SEX == 1, "Man", "Woman"),
    Approval = if_else(POL1DT_W52 == 1, "Approve", "Disapprove"),
    Age = if_else(
      F_AGECAT == 1, "18-29",
      if_else(F_AGECAT == 2, "30-49",
        if_else(F_AGECAT == 3, "50-64",
          if_else(F_AGECAT == 4, "65+", "N/A")))))
```





Trump Approval Rating
---

President Trump led a tumultuous presidency, where approval ratings varied greatly. In the data collected through Pew from July 2019, results fall along party lines. nearly 80% of both Republican men and women supported President Trump, while nearly 75% Democrat men and over 75% of Democrat women disapproved of the President. 

```{r}
##assigning vars and demonstrating Trump approval 
trumpvar <-
  var2 |>
  group_by(Party, Gender) |>
  count(Approval) |>
  mutate(pct = n / sum(n) *100 ) |>
  select(Party, Gender, Approval, pct)

##plotting Trump approval 
trumpvar |>
  ggplot(mapping = aes(x = Approval, y = pct, fill = Party)) +
  geom_col(position = "dodge") +
  facet_grid(rows = vars(Gender)) +
  scale_fill_manual(values = c(Democrat = "steelblue", Republican = "indianred")) +
  labs(title = "Approval Ratings of President Donald Trump",
  x = "Approval",
  y = "Percentage",
  subtitle = "POL1DT_W52, ATP W52, Pew Research Center") 

##printing table of Trump approval data
knitr::kable(trumpvar, col.names = c("Political Party", "Gender", "Approval Status", "Percent"),
             caption = "Approval Ratings of President Donald Trump")

```

Perspectives on the Economy Under Trump
---


Pew's report on the current economic climate under Trump is more surprising, as results vary among the  two primary parties greatly. Men that identify as either Republican or Democrat largely agreed, rating  the current economy as "Good", serving as the second best response presented. Women disagreed slightly more, with Republican women largely responding "Excellent" and "Good" with Democrats averaging nearer    to the "Only Fair" rating response. 


```{r}
##filtering and assigning data for Trump current econ perspectives 
var_trump <-
  var2 |>
  filter(ECON1_W52 == 1 | ECON1_W52 == 2 | ECON1_W52 == 3 | ECON1_W52 == 4 | ECON1_W52 == 5) |>
  mutate(ECON1_W52 = if_else(ECON1_W52 == 1, "Excellent",
  if_else(ECON1_W52 == 2, "Good",
  if_else(ECON1_W52 == 3, "Only Fair",
  if_else(ECON1_W52 == 4, "Poor", "No Change"))))) |>
  group_by(Party, Gender) |>
  count(ECON1_W52) |>
  mutate(pct = n / sum(n) *100 ) |>
  select(Party, Gender, ECON1_W52, pct)

##assigning linear regression for Trump current econ perspectives
model_trump <- lm(pct ~ ECON1_W52, data = var_trump)

##filtering and assigning data for Trump future econ perspectives 
var_trump2 <-
  var2 |>
  filter(ECON1B_W52 == 1 | ECON1B_W52 == 2 | ECON1B_W52 == 3 | ECON1B_W52 == 4) |>
  mutate(ECON1B_W52 = if_else(ECON1B_W52 == 1, "Better",
  if_else(ECON1B_W52 == 2, "Worse",          
  if_else(ECON1B_W52 == 3, "About Same", "Unsure")))) |>
  group_by(Party, Gender) |>
  count(ECON1B_W52) |>
  mutate(pct = n / sum(n) *100 ) |>
  select(Party, Gender, ECON1B_W52, pct)

##assigning linear regression for Trump future econ perspectives
model_trump2 <- lm(pct ~ ECON1B_W52, data = var_trump2)


##plotting data for Trump current econ perspectives
trump_plot1 <- 
  var_trump |>
  ggplot(mapping = aes(x = ECON1_W52, y = pct, fill = Party)) +
  geom_col(position = "dodge") +
  facet_grid(rows = vars(Gender)) +
  scale_fill_manual(values = c(Democrat = "steelblue", Republican = "indianred")) +
  labs(title = "Thoughts on Current Economy Under Trump",
  x = "Feelings Towards Economy",
  y = "Percentage", 
  subtitle = "ECON1_W52, ATP W52, Pew Research Center") 

##plotting data for Trump future econ perspectives
trump_plot2 <- 
  var_trump2 |>
  ggplot(mapping = aes(x = ECON1B_W52, y = pct, fill = Party)) +
  geom_col(position = "dodge") +
  facet_grid(rows = vars(Gender)) +
  scale_fill_manual(values = c(Democrat = "steelblue", Republican = "indianred")) +
  labs(title = "Thoughts on Future Economy Under Trump",
  x = "Feelings Towards Future Economy",
  y = "Percentage",
  subtitle = "ECON1B_W52, ATP W52, Pew Research Center") 

##printing plot and table for Trump current econ perspectives
trump_plot1
knitr::kable(var_trump, col.names = c("Party", "Gender", "Econ Feelings", "Percentage"),
             caption = "Thoughts on Current Economy Under Trump")

```


For this demonstration of linear regression, I have identified `pct` as the dependent variable, depending on the independent variable, which is `ECON1_W52`. The `pct` intercept value of 20.48 demonstrates holding all else equal, what the `pct` approval rating would equal. Then, taking the responses to `ECON1_W52` into account, we find that `ECON1_W52Good` has an intercept of 26.75, meaning for each unit increase in average response of `ECON1_W52Good`, there will be a nearly 27 percentage point increase in approval rating. Continuing, the intercept for `ECON1_W52OnlyFair` equals 5.05, meaning there is a smaller increase in approval of President Trump for those who rank the economy as relatively middle of the road. For those who identify the economy as poor, there is a decrease of 13.71 percentage points, indicating that those who find the economy to be negative are inclined to support President Trump almost 14 percentage points less than other respondents. This reinforces the hypothesis, that respondents who align with the party of the President are more likely to believe the economy is working well. 



```{r}
##printing table of linear regression for Trump current econ perspectives
modelsummary::modelsummary(model_trump,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           caption = "Regression on Thoughts on Current Economy Under Trump")
```


When asked about one year from the date of polling, respondents were generally quite optimistic under President Trump. These results come from pre-Covid 19, meaning that almost all of the participants were unfortunately quite wrong. In general, Republicans responded resoundingly in the positive, claiming the economy would either be the same or better in one calendar year. Democrats responded primarily in the "About Same" category, with an unsurprising showing in the "Worse" response category. 

```{r}
##printing plot and table for Trump future econ perspectives
trump_plot2
knitr::kable(var_trump2, col.names = c("Party", "Gender", "Future Econ Feelings", "Percentage"),
             caption = "Thoughts on Future Economy Under Trump")
```


In the below linear regression, `pct` serves as the dependent variable, with `ECON1B_W52` as the independent variable. First, the intercept, which equals 41.491, signifies the average predicted value of "pct" sans responses from `ECON1B_W52Better` and `ECON1B_W52Worse`. Next, the coefficient for `ECON1B_W52Better` equals -4.65, demonstrating that holding all others constant, respondents who claim the future economy will be better on average decrease their `pct` value, which demonstrates their approval rating of President Trump, by 4.65 points. Accordingly, the result for those who responded `ECON1B_W52Worse` experienced a -19.83 percentage point change in `pct` approval of President Trump. This reinforces the idea that those who support the President find the economy to be stronger than those who do not support the President. The intercept demonstrates a significant p-value, equaling less than .001. Neither `ECON1B_W52Better` nor `ECON1B_W52Worse` demonstrate significant p-values with a confidence level of .05. 

```{r}
##printing table of linear regression for Trump future econ perspective
modelsummary::modelsummary(model_trump2,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"),
                           gof_map = c("nobs", "r.squared", "adj.r.squared"),
                           caption = "Regression on Thoughts on Future Economy Under Trump")
```

To conclude the linear regression portion, I ran a multi-variable regression analyzing multiple factors prevalent to the approval rating, `pct`, of President Joe Biden. Factors include `ECON1_W52`,     `ECON1B_W52`, `Gender` and `Party`. The evidence demonstrates that the two most statistically significant values come from respondents of `ECON1_W52Good` and `ECON1B_W52Unsure`, where values are impressively negative. This demonstrates that in `ECON1_W52`, those that rated the current economy as good are, on average, experiencing an increase by point of 8.42 percentage points, meaning they experience a substantial increase in approval rating response. Those in the `ECON1B_W52` category of "Unsure" were not aware of the future state of the economy, and experienced a strong, -15.2 percent point change in approval. Both of these variables have significant p-values, with a confidence variable of .05. With p-values under .05 and .001 respectively, the data demonstrates significance. Interestingly, the `Gender` and `Party` variables serve little significance, with very minor negative and positive values, respectively. 


``` {r}
##assigning, filtering, adjusting data for multivariable regression
finalvar <-
  var2 |>
  filter(ECON1_W52 == 1 | ECON1_W52 == 2 | ECON1_W52 == 3 | ECON1_W52 == 4 | ECON1_W52 == 5 | 
  ECON1B_W52 == 1 | ECON1B_W52 == 2 | ECON1B_W52 == 3 | ECON1B_W52 == 4) |>
  mutate(ECON1_W52 = if_else(ECON1_W52 == 1, "Excellent",
  if_else(ECON1_W52 == 2, "Good",
  if_else(ECON1_W52 == 3, "Only Fair",
  if_else(ECON1_W52 == 4, "Poor", "No Change"))))) |>
  mutate(ECON1B_W52 = if_else(ECON1B_W52 == 1, "Better",
  if_else(ECON1B_W52 == 2, "Worse",          
  if_else(ECON1B_W52 == 3, "About Same", "Unsure")))) |>
  group_by(Party, Gender) |>
  count(ECON1_W52, ECON1B_W52) |>
  mutate(pct = n / sum(n) *100 ) |>
  select(pct, Party, Gender, ECON1_W52, ECON1B_W52)
##assigning linear regression for multivariable demonstration
finalvar <- lm(pct ~ ECON1_W52 + ECON1B_W52 + Gender + Party, data = finalvar)

##assigning names of list for table of multivariable demonstration
names <- list("ECON1_W52Good" = "Current - Good",
              "ECON1_W52No Change" = "Current - No Change",
              "ECON1_W52Only Fair" = "Current - Only Fair",
              "ECON1_W52Poor" = "Current - Poor",
              "ECON1B_W52Better" = "Future - Better",
              "ECON1B_W52Unsure" = "Future - Unsure",
              "ECON1B_W52Worse" = "Future - Worse",
              "GenderWoman" = "Gender: Woman",
              "PartyRepublican" = "Party: Republican")

##printing table of multivariable regression demonstration
modelsummary::modelsummary(finalvar,
                           statistic = c("s.e. = {std.error}",
                                         "p = {p.value}"), 
                           coef_map = names,
                           gof_map = c("nobs", "r.squared", "adj.r.squared"), 
                           title = "Regression on Multiple Factors Impacting Approval Ratings")
```





Conclusion
---
Political affairs consistently demonstrate incredible levels of uncertainty, confusion, and general mistrust that are rarely seen in other situations. When analyzing such substantial metrics, like Presidential approval ratings, and speculation on the economy, it is important to get things right. Unfortunately, that is not possible within the scope of this project. Just with the approval rating, countless confounding variables could have a substantial impact on the metric. For example, do men like Joe Biden less because of a comment he made on gun rights the week of the survey, or do they dislike his suit color? Do women believe the economy is doing badly because the stocks they trade in their jobs took a dip today, or maybe because they like to drink more expensive milk? Especially with the last few administrations, approval ratings vary hour by hour. With the utilization of social media, news of a slipup, comment, or thought is spread so quickly that people change their opinions multiple times during a news cycle. 

On the topic of the hypothesis, it appears economic approval and presidential approval line up more with the Trump presidency than Biden's. This, again, is likely due to issues that were not investigated here. The data on President Trump was collected before Covid-19, which is well known to have changed the course of the international economy permanently. President Trump was also very well known as an active member of social media where his comments would raise and lower markets across the world. Some Americans fervently supported Trump, while others despised him.

In the case of Joe Biden, his economic prowess is also not abundantly clear. In a post-Covid world, one can easily speculate that Biden was 'dealt a bad hand' and polled Americans are experiencing fear of another recession-type event. Americans have paid more attention to the economy than in previous years, and Covid has a lot to do with it. 

Finally, the hypothesis, which states that when a President is in office, those who align with the POTUS' party will believe the economy is strong is seemingly disproven. No clear or resounding information demonstrates a consistent relationship between economic climate and approval ratings. The President does more than influence the economy, and this report reiterates that important point. 

If presented with more time and resources, I would utilize more data sets with similar questions. The Pew Research Center strives to establish relationships between surveys, but it is not always viable to ask every question they want to ask. For this reason, I was only able to utilize two data sets. This would allow me to compare trends across a greater period, which would likely allow me to minimize the substantial impact Covid had on the data collected. 


